---
title: "Feature Selection"
output: pdf_document
---


**Aim**
To generate a classifier to classify breast tumours into Integrative Clusters (iClust) using data obtained using NanoString platform.


#Summary of Samples
480 samples with known iClust status
212 genes selected previously from the original classifier


```{r Preparation, echo=FALSE}
#loading libraries
library(ggplot2)
library(lattice)
library(caret)
library(foreach)
library(iterators)
#library(doMC)
library(parallel)
library(rpart)

#set up multicore
#registerDoMC(cores = 4)

#loading data
setwd("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/R scripts")
load("nanoStringIC_DF.RData")

#remove na
nanoStringIC_DF <- na.omit(nanoStringIC_DF)

```

#Feature Selection

This step is to select genes that are sufficiently differentially expressed between the groups.

All of the 212 iClust genes are included.
Approximately 100 immune genes are also differentially expressed.  However these are not included here.  They will be included later.


```{r Feature Selection, echo=FALSE}
features <- data.frame(
  p=sapply(2:ncol(nanoStringIC_DF),function(i){
    anova(lm(log(nanoStringIC_DF[,i]+0.001) ~ nanoStringIC_DF$iClust))$"Pr(>F)"[1]
  }),
  ID=colnames(nanoStringIC_DF)[-1]
)
feature.ordered <- features[order(features$p),]
feature.ordered
```