Immune Genes using Nanostring


```{r}
#loading library
library(ggplot2)
library(NanoStringNorm)
library(reshape2)
library(ggplot2)
library(lattice)
library(caret)
library(foreach)
library(iterators)
library(parallel)
library(doMC)
library(randomForest)
registerDoMC(cores = 4)
```


```{r}
#loading data

batchNumber <- "Combined"
metadataFolder <- "/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Other Info/"
nanostringDataFolder <- paste0("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Digital Analyser/",batchNumber,"/")
outputFolder <- paste0("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Nanostring QC/",batchNumber,"QC/")


#Get int clust data
iClust <- read.csv(paste0(metadataFolder,"IntClust.csv"))


#Get clinical data
clinical <- read.csv(paste0(metadataFolder,"Metabric.Clinical.csv"))
colnames(clinical)[2] <- "ID"

#getting MB-AD files
nanoStringFiles <- list.files(nanostringDataFolder, pattern = "MB-AD", recursive = T, full.names = T)
#Getting NT files
nanoStringFiles<-c(nanoStringFiles, list.files(nanostringDataFolder, pattern="NT", recursive = T, full.names=T))
#Getting TN files
nanoStringFiles<-c(nanoStringFiles,
                   list.files(nanostringDataFolder, pattern="TN", recursive = T, full.names=T))

#get csv
nanoStringData <- lapply(nanoStringFiles, read.csv,
                         skip = 26, header = T, nrows = 773,quote=NULL,
                         stringsAsFactors = F)

#extract id from file id
SampleID <- regmatches(nanoStringFiles, regexpr("MB-AD-[0-9]*", nanoStringFiles))
SampleID <- c(SampleID, regmatches(nanoStringFiles, regexpr("NT-[0-9]*", nanoStringFiles)))
SampleID <- c(SampleID, regmatches(nanoStringFiles, regexpr("TN-[0-9]*", nanoStringFiles)))


for (i in seq_along(nanoStringData)){
  colnames(nanoStringData[[i]])[colnames(nanoStringData[[i]])=="Count"] <-SampleID[[i]]
}

#convert list to dataframe
nanoStringData <- do.call("cbind", nanoStringData)
nanoStringData <- nanoStringData[, !duplicated(colnames(nanoStringData), fromLast = FALSE)] 



#normalisation using NanoStringNorm
normalisedNanoStringData <- NanoStringNorm(
  x = nanoStringData,
  anno = NA,
  CodeCount = 'sum',
  Background = 'mean',
  SampleContent = 'housekeeping.sum',
  round.values = FALSE,
  take.log = FALSE,
  return.matrix.of.endogenous.probes = TRUE
)
```




nanostring norm suggested removing some samples ? do or not?

```{r}
#remove ones suggested by nanostring norm


```

combine iclust with nanostring data

```{r}
#data tidied up
normData <- t(normalisedNanoStringData)
normData <- as.data.frame(normData)
normData$ID <- rownames(normData)
normData$ID <- sub("-AD-", "-", normData$ID, fixed = T)
normData <- merge(normData, iClust, by="ID", all.x=T)
normData <- merge(normData, clinical, by="ID", all.x=T)
normData <- normData[!is.na(normData$IntClust), ]
```



do not do na.omit otherwises will lose a lot of data at this stage



feature selection wrt IntClust
```{r}

features <- data.frame(
  p=sapply(2:(ncol(normData)-51),function(i){
    anova(lm(log(normData[,i]+0.001) ~ normData$IntClust))$"Pr(>F)"[1]
  }),
  ID=colnames(normData)[2:(ncol(normData)-51)]
)
```

select features that have p<0.05
```{r}
featured <- features[which(features$p<0.05),]
normData_featured <- cbind(normData$ID, normData$IntClust,normData[, which(colnames(normData) %in% featured$ID)])
colnames(normData_featured)[1:2] <- c("ID", "IntClust")
rownames(normData_featured) <- normData_featured$ID
normData_featured$ID <- NULL
colnames(normData_featured) <- gsub(x = colnames(normData_featured), pattern = "-", replacement = "\\.")



```

try random forest
```{r}
nanostring_rf <- randomForest(IntClust~., data=normData_featured, ntree=200)
print(nanostring_rf)
```

```{r}
normData_featured$IntClust<-paste0("x",normData_featured$IntClust)
ctrl <- trainControl(method = "repeatedcv", savePred=T, classProb=T, repeats=100)
mod <- train(IntClust~., data=normData_featured, method = "rf", trControl = ctrl)
mod$finalModel$confusion
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

