---
title: "using marcus' data and to see whether sample selection has problem or probe selection"
output: pdf_document
---
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#loading data
load("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/R scripts/220features.RData")
library(iC10)
library(ggplot2)
library(reshape2)
library(data.table)

#save(IntClust, file="IntClust.RData")
load("IntClust.RData")
load("nanoStringIC_DF.RData")
```

Firstly using Marcus' data, run the 220 genes, 1980 samples with iC10 package to see what the accuracy is
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
features <- matchFeatures(Exp=t(DATA), Exp.by.feat="gene")
features <- normalizeFeatures(features, "scale")
res <- iC10(features)

predicted <- data.frame(res$class)
rownames(predicted) <- sub("-", ".", fixed = TRUE, rownames(predicted))
#dim(predicted)
rownames(IntClust) <- IntClust$ID
IntClust <- IntClust[which(IntClust$ID %in% rownames(predicted)), ]
IntClust <- IntClust[order(match(IntClust$ID, rownames(predicted))), ]
#sum(IntClust$ID==rownames(predicted))

#accuracy
ic10_marcus_table <- table(predicted$res.class, IntClust$IntClust)
sum(diag(ic10_marcus_table))/sum(ic10_marcus_table)

#0.8345287

```

Now we use the samples that were used for the NS

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#rownames(nanoStringIC_DF)

DATA_nssamples <- DATA[which(rownames(DATA) %in% rownames(nanoStringIC_DF)), ]
#dim(DATA_nssamples)
#456 metabric samples, 220 genes)

#now run the iC10
features <- matchFeatures(Exp=t(DATA_nssamples), Exp.by.feat="gene")
features <- normalizeFeatures(features, "scale")
res <- iC10(features)

predicted_nssamples <- data.frame(res$class)
rownames(predicted_nssamples) <- sub("-", ".", fixed = TRUE, rownames(predicted_nssamples))
#dim(predicted_nssamples)
IntClust_nssamples <- IntClust[which(IntClust$ID %in% rownames(predicted_nssamples)), ]
IntClust_nssamples <- IntClust_nssamples[order(match(IntClust_nssamples$ID, rownames(predicted_nssamples))), ]
#sum(IntClust_nssamples$ID==rownames(predicted_nssamples))

#accuracy
ic10_marcus_table_nssamples <- table(predicted_nssamples$res.class, IntClust_nssamples$IntClust)
sum(diag(ic10_marcus_table_nssamples))/sum(ic10_marcus_table_nssamples)

#0.8442982
```


now with nanostring Data

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
nanostring_metabric <- nanoStringIC_DF[1:458, ]
rownames(nanostring_metabric) <- sub("-", ".", fixed = TRUE, rownames(nanostring_metabric))
nanostring_metabric <- nanostring_metabric[which(rownames(nanostring_metabric) %in% IntClust$ID), ]
rownames(nanostring_metabric) <- sub("-", ".", fixed = TRUE, rownames(nanostring_metabric))

features_ns <- matchFeatures(Exp=t(nanostring_metabric[, -1]), Exp.by.feat="gene")
features_ns <- normalizeFeatures(features_ns, "scale")
res <- iC10(features_ns)

predicted_ns <- data.frame(res$class)

#dim(predicted_ns)
rownames(predicted_ns) <- sub("-", ".", fixed = TRUE, rownames(predicted_ns))
IntClust_ns <- IntClust[which(IntClust$ID %in% rownames(predicted_ns)), ]

IntClust_ns <- IntClust_ns[order(match(IntClust_ns$ID, rownames(predicted_ns))), ]
#mean(IntClust_ns$ID==rownames(predicted_ns))

correct_prediction <- data.frame(cbind(IntClust_ns$ID, IntClust_ns$IntClust, predicted_ns$res.class))
names(correct_prediction) <- c('MBID', 'IntClust', "predictedIntClust")
correct_prediction <- correct_prediction[which(correct_prediction$IntClust==correct_prediction$predictedIntClust), ]
write.csv(correct_prediction, file="correct_prediction.csv")

#accuracy
ic10_marcus_table_nsdata <- table(predicted_ns$res.class, IntClust_ns$IntClust)
sum(diag(ic10_marcus_table_nsdata))/sum(ic10_marcus_table_nsdata)


#0.4868421
```

now there are 3 genes difference between the nanostring data and the illumina data with same samples.  we now limit the illumina one to the same 139 genes and rerun, to make sure these are not crucial genes

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

#using DATA_nssamples
#456 metabric samples, 220 genes)
#now remove the genes
extra3genes <- rownames(features$Exp)[-which(rownames(features$Exp) %in% rownames(features_ns$Exp))]

DATA_nssamples_139genes <- DATA_nssamples[, -which(colnames(DATA_nssamples) %in% extra3genes)]


#now run the iC10
features_ns139genes <- matchFeatures(Exp=t(DATA_nssamples_139genes), Exp.by.feat="gene")
features_ns139genes <- normalizeFeatures(features_ns139genes, "scale")
res <- iC10(features_ns139genes)

predicted_nssamples_139genes <- data.frame(res$class)
rownames(predicted_nssamples_139genes) <- sub("-", ".", fixed = TRUE, rownames(predicted_nssamples_139genes))
#dim(predicted_nssamples_139genes)
IntClust_nssamples_139genes <- IntClust[which(IntClust$ID %in% rownames(predicted_nssamples_139genes)), ]
IntClust_nssamples_139genes <- IntClust_nssamples_139genes[order(match(IntClust_nssamples_139genes$ID, rownames(predicted_nssamples_139genes))), ]
#sum(IntClust_nssamples_139genes$ID==rownames(predicted_nssamples_139genes))

#accuracy
ic10_marcus_table_nssamples_139genes <- table(predicted_nssamples_139genes$res.class, IntClust_nssamples_139genes$IntClust)
sum(diag(ic10_marcus_table_nssamples_139genes))/sum(ic10_marcus_table_nssamples_139genes)

#0.8552632
```

ok then in that case the correlation of genes are not good. get the correlations

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#comparing correlation
#save(cors_gene, file="cors_gene.RData")
load("cors_gene.RData")

cors_gene <- cors_gene[order(cors_gene$cors), ]

cors_gene_139 <- cors_gene[which(cors_gene$Genes %in% rownames(features$Exp)), ]

ggplot(cors_gene_139, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Data")


```


Now we sort the genes in order of correlation (high to low), and take the first 50 genes and compare the accuracy of prediction.  

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rownames(cors_gene_139) <- NULL
cors_gene_139 <- cors_gene_139[order(-cors_gene_139$cors), ]

#now subset the illumina and nanostring data using the same order of genes for subsetting
illumina_50corgenes <- DATA_nssamples_139genes[ , which(colnames(DATA_nssamples_139genes) %in% cors_gene_139$Genes[1:50])]
nanostring_50corgenes <- nanostring_metabric[, which(colnames(nanostring_metabric) %in% cors_gene_139$Genes[1:50])]

#now run the iC10 - firstly illumina data
features_50genes <- matchFeatures(Exp=t(illumina_50corgenes), Exp.by.feat="gene")
features_50genes <- normalizeFeatures(features_50genes, "scale")
res <- iC10(features_50genes)

predicted_ill_50genes <- data.frame(res$class)
rownames(predicted_ill_50genes) <- sub("-", ".", fixed = TRUE, rownames(predicted_ill_50genes))
#dim(predicted_ill_50genes)
IntClust_ill_50genes <- IntClust[which(IntClust$ID %in% rownames(predicted_ill_50genes)), ]
IntClust_ill_50genes <- IntClust_ill_50genes[order(match(IntClust_ill_50genes$ID, rownames(predicted_ill_50genes))), ]
#mean(IntClust_ill_50genes$ID==rownames(predicted_ill_50genes))

#accuracy
ic10_marcus_table_50genes <- table(predicted_ill_50genes$res.class, IntClust_ill_50genes$IntClust)
print("Illumina:")
sum(diag(ic10_marcus_table_50genes))/sum(ic10_marcus_table_50genes)

#0.7258772
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#now run the iC10 - nanostring data
features_50genes <- matchFeatures(Exp=t(nanostring_50corgenes), Exp.by.feat="gene")
features_50genes <- normalizeFeatures(features_50genes, "scale")
res <- iC10(features_50genes)

predicted_ns_50genes <- data.frame(res$class)
rownames(predicted_ns_50genes) <- sub("-", ".", fixed = TRUE, rownames(predicted_ns_50genes))
#dim(predicted_ns_50genes)
IntClust_ns_50genes <- IntClust[which(IntClust$ID %in% rownames(predicted_ns_50genes)), ]
IntClust_ns_50genes <- IntClust_ns_50genes[order(match(IntClust_ns_50genes$ID, rownames(predicted_ns_50genes))), ]
#mean(IntClust_ns_50genes$ID==rownames(predicted_ns_50genes))

#accuracy
ic10_marcus_table_50genes <- table(predicted_ns_50genes$res.class, IntClust_ns_50genes$IntClust)
print("NanoString:")
sum(diag(ic10_marcus_table_50genes))/sum(ic10_marcus_table_50genes)

#0.4451754
```

It would appear that the difference has already occured within this 50 genes.  Now try with 25

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rownames(cors_gene_139) <- NULL
cors_gene_139 <- cors_gene_139[order(-cors_gene_139$cors), ]

#now subset the illumina and nanostring data using the same order of genes for subsetting
illumina_25corgenes <- DATA_nssamples_139genes[ , which(colnames(DATA_nssamples_139genes) %in% cors_gene_139$Genes[1:25])]
nanostring_25corgenes <- nanostring_metabric[, which(colnames(nanostring_metabric) %in% cors_gene_139$Genes[1:25])]

#now run the iC10 - firstly illumina data
features_25genes <- matchFeatures(Exp=t(illumina_25corgenes), Exp.by.feat="gene")
features_25genes <- normalizeFeatures(features_25genes, "scale")
res <- iC10(features_25genes)

predicted_ill_25genes <- data.frame(res$class)
rownames(predicted_ill_25genes) <- sub("-", ".", fixed = TRUE, rownames(predicted_ill_25genes))
#dim(predicted_ill_25genes)
IntClust_ill_25genes <- IntClust[which(IntClust$ID %in% rownames(predicted_ill_25genes)), ]
IntClust_ill_25genes <- IntClust_ill_25genes[order(match(IntClust_ill_25genes$ID, rownames(predicted_ill_25genes))), ]
#mean(IntClust_ill_25genes$ID==rownames(predicted_ill_25genes))

#accuracy
ic10_marcus_table_25genes <- table(predicted_ill_25genes$res.class, IntClust_ill_25genes$IntClust)
print("Illumina:")
sum(diag(ic10_marcus_table_25genes))/sum(ic10_marcus_table_25genes)

#0.6491228


#now run the iC10 - nanostring data
features_25genes <- matchFeatures(Exp=t(nanostring_25corgenes), Exp.by.feat="gene")
features_25genes <- normalizeFeatures(features_25genes, "scale")
res <- iC10(features_25genes)

predicted_ns_25genes <- data.frame(res$class)
rownames(predicted_ns_25genes) <- sub("-", ".", fixed = TRUE, rownames(predicted_ns_25genes))
#dim(predicted_ns_25genes)
IntClust_ns_25genes <- IntClust[which(IntClust$ID %in% rownames(predicted_ns_25genes)), ]
IntClust_ns_25genes <- IntClust_ns_25genes[order(match(IntClust_ns_25genes$ID, rownames(predicted_ns_25genes))), ]
#mean(IntClust_ns_25genes$ID==rownames(predicted_ns_25genes))

#accuracy
ic10_marcus_table_25genes <- table(predicted_ns_25genes$res.class, IntClust_ns_25genes$IntClust)
print("NanoString:")
sum(diag(ic10_marcus_table_25genes))/sum(ic10_marcus_table_25genes)

#0.44098
```

Now to see how the genes correlates within themselves
Take RSF1 for example, find the gene that is most correlated with RSF1, which is CLNS1A.  
Left is using Illumina data
Right is using Nanostring Data

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#illumina sample:
#DATA_nssamples_139genes
nanostring_metabric_139genes <- nanostring_metabric[, which(colnames(nanostring_metabric) %in% colnames(DATA_nssamples_139genes))]

#easier to compare all 217..... as otherwise can't find the alternative names

cor_illumina <- cor(DATA_nssamples_139genes)
cor_nanostring <- cor(nanostring_metabric_139genes)

#find out which ones are the most correlated with RSF1
#names(sort(cor_illumina["RSF1", ]))
#see how well the data for nanostring correlate with illumina data
par(mfrow=c(1,2))
plot(DATA[, "RSF1"], DATA[, "CLNS1A" ])
plot(nanostring_metabric[, "RSF1"], nanostring_metabric[, "CLNS1A" ])

melted_cor_illumina <- melt(cor_illumina)
cor_illumina_139 <- cor(DATA_nssamples_139genes[, which(colnames(DATA_nssamples_139genes) %in% rownames(features_ns))])
melted_cor_illumina_139 <- melt(cor_illumina_139)

#heatmap for genes correlating with each other
#ggplot(melted_cor_illumina, aes(x=Var1, y=Var2, fill=value))+
#  geom_tile()

```

We now plot RSF1 in illumina vs nanostring for each integrative clusters

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#make a RSF1 dataframe
nanoStringIC_DF <- na.omit(nanoStringIC_DF)
nanostring_metabric <- nanoStringIC_DF[1:458,]
#rownames(nanostring_metabric)
#rownames(DATA_nssamples)
#DATA_nssamples <- DATA[which(rownames(DATA) %in% rownames(nanostring_metabric)), ]
nanostring_metabric <- nanostring_metabric[which(rownames(nanostring_metabric) %in% rownames(DATA_nssamples)), ]

#dim(DATA_nssamples)
#dim(nanostring_metabric)
nanostring_metabric <- nanostring_metabric[order(match(rownames(nanostring_metabric),  rownames(DATA_nssamples))), ]




RSF1_df <- data.frame(cbind(nanostring_metabric[,1], nanostring_metabric[, "RSF1"], DATA_nssamples[, "RSF1"]))

colnames(RSF1_df) <- c("iClust", "NanoString", "Illumina")

ggplot(RSF1_df, aes(x=log(NanoString), y=Illumina))+
  geom_point()+
  facet_wrap(~iClust)+
  theme_bw()+
  ggtitle("RSF1 gene expression, Illumina vs NanoString for each iClust")+
  geom_text(aes(label=rownames(RSF1_df)), size=2, nudge_y = +0.15)



```


We then identified the outliers, and see if these are due to poor correlation - which isn't

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#who are the outliers?
#firstly sort out a dataframe that has the right gene names
DATA_nssamples_212genes <- DATA_nssamples[, which(colnames(DATA_nssamples) %in% colnames(nanostring_metabric))]
DATA_nssamples_212genes <- DATA_nssamples_212genes[, order(match(colnames(DATA_nssamples_212genes), colnames(nanostring_metabric)[-1]))]
#mean(colnames(DATA_nssamples_212genes) == colnames(nanostring_metabric)[-1])

outliers_mbid <- c("MB-0428", "MB-0228")
outliers_illumina <- DATA_nssamples_212genes[outliers_mbid, ]
outliers_nanostring <- as.matrix(nanostring_metabric[outliers_mbid, -1])
#mean(colnames(outliers_illumina)==colnames(outliers_nanostring))
cor_outliers <- sapply(1:nrow(outliers_illumina), function (x) cor(outliers_illumina[x,], outliers_nanostring[x,]))

par(mfrow=c(1,2))
for (i in 1:2){
  plot(outliers_illumina[i, ], log(outliers_nanostring[i,]), pch=20)
  text(outliers_illumina[i, ], log(outliers_nanostring[i,])+0.2, colnames(outliers_illumina), cex=0.6)
}

cor_outliers
# colnames(outliers_illumina)
#, type="n"
```



```{r eval=FALSE, include=FALSE}
#Now we look at each sample that should be in iC2 which are not correctly classified


features_ic2 <- matchFeatures(Exp=t(nanostring_metabric[,-1]), Exp.by.feat="gene")
features_ic2 <- normalizeFeatures(features_ic2, "scale")
res <- iC10(features_ic2)

predicted_ic2 <- data.frame(res$class)
#mean(rownames(predicted_ic2)==rownames(nanostring_metabric))
predicted_ic2$original <- nanostring_metabric$iClust
#now subset it to those that are meant to be in iC2 but is predicted wrongly
predicted_ic2_1 <- predicted_ic2[which((predicted_ic2$original==2) & (predicted_ic2$res.class != 2)), ]
#now subset it to those that are NOT meant to be in iC2 but is predicted to be iC2
predicted_ic2_2 <- predicted_ic2[which((predicted_ic2$original!=2) & (predicted_ic2$res.class == 2)), ]

#now plot rsf1 and clns1a in these samples
#firstly the ones that shoudl have been in ic2
par(mfrow=c(1,2))
plot(DATA[c(rownames(predicted_ic2_1)), "RSF1"], DATA[c(rownames(predicted_ic2_1)), "CLNS1A" ])
plot(nanostring_metabric[c(rownames(predicted_ic2_1)), "RSF1"], nanostring_metabric[c(rownames(predicted_ic2_1)), "CLNS1A" ])

#now the ones that are wrongly classified into iC2
par(mfrow=c(1,2))
plot(DATA[c(rownames(predicted_ic2_2)), "RSF1"], DATA[c(rownames(predicted_ic2_2)), "CLNS1A" ])
plot(nanostring_metabric[c(rownames(predicted_ic2_2)), "RSF1"], nanostring_metabric[c(rownames(predicted_ic2_2)), "CLNS1A" ])


#now plot for each of these incorrectly classified samples, the expression value for each gene and their value
predicted_ic2_both <- data.frame(rbind(predicted_ic2_1, predicted_ic2_2))
wrong_ic2 <- rbind(data.table(DATA_nssamples_212genes[rownames(predicted_ic2_both),], keep.rownames = T), data.table(log(nanostring_metabric[rownames(predicted_ic2_both),-1]), keep.rownames = T) )
wrong_ic2$Method <- c( rep("Illumina", 11), rep("NanoString", 11))
wrong_ic2_melted <- melt(wrong_ic2)

ggplot(wrong_ic2_melted, aes(x=variable, y=value))+
  geom_point()+
  coord_flip()+
  facet_grid(rn ~ Method, scales = "free_x")

samples_list <- rownames(predicted_ic2_both)

pdf("wrong_ic2.PDF")
par(mfrow=c(1,1))
par(mar = rep(2, 4))
for (i in 1:11) {
  tmp <- wrong_ic2_melted[rn==samples_list[i], ]
  tmp_i <- tmp[Method=="Illumina", ]
  tmp_i <- tmp_i[order(tmp_i$value), ]
  tmp_n <- tmp[Method=="NanoString", ]
  tmp_n <- tmp_n[order(tmp_n$value), ]
  dotchart(tmp_i$value, labels=tmp_i$variable, cex = 0.5)
  dotchart(tmp_n$value, labels=tmp_n$variable, cex=0.5)
}
dev.off()

pdf("wrong_ic2_list.PDF")
for (i in 1:11) {
  tmp <- wrong_ic2_melted[rn==samples_list[i], ]
  tmp_i <- tmp[Method=="Illumina", ]
  tmp_i <- tmp_i[order(tmp_i$value), ]
  tmp_n <- tmp[Method=="NanoString", ]
  tmp_n <- tmp_n[order(tmp_n$value), ]
  print(samples_list[i])
  print("high")
  print(tmp_i$variable[c(1:10, 203:212)])
  print("low")
  print(tmp_n$variable[c(1:10, 203:212)])
}
dev.off()






```

We now look at outliers for each sample

```{r}
nanostring_metabric <- nanostring_metabric[order(match(rownames(nanostring_metabric),  rownames(DATA_nssamples))), ]

#sum(rownames(nanostring_metabric)==rownames(DATA_nssamples))
DATA_nssamples_matched <- DATA_nssamples[, which(colnames(DATA_nssamples) %in% colnames(nanostring_metabric))]
DATA_nssamples_matched <- DATA_nssamples_matched[, order(match(colnames(DATA_nssamples_matched), colnames(nanostring_metabric)[-1]))]
#sum(colnames(nanostring_metabric)[-1] == colnames(DATA_nssamples_matched))
library(ggplot2)
gene_per_cluster <- list()
pdf("Plots of Each Gene per Cluster.PDF")
for (i in 2:ncol(nanostring_metabric)){
  tmp.name <- colnames(nanostring_metabric)[i]
  gene_per_cluster[[i-1]] <- data.frame(cbind(nanostring_metabric[,1], nanostring_metabric[, i], DATA_nssamples[, i-1]))
  colnames(gene_per_cluster[[i-1]]) <- c("iClust", "NanoString", "Illumina")
  plot<- ggplot(gene_per_cluster[[i-1]], aes(x=log(NanoString), y=Illumina))+
          geom_point()+
          facet_wrap(~iClust)+
          theme_bw()+
        ggtitle(paste0(tmp.name, "gene expression, Illumina vs NanoString for each iClust"))+
          geom_text(aes(label=rownames(gene_per_cluster[[i-1]])), size=2, nudge_y = +0.15)
print(plot)  
}
dev.off()

outlier <- data.frame(matrix(,ncol = 212, nrow = 100))
for (i in 1:length(gene_per_cluster)){
  tmp.df <- gene_per_cluster[[i]]
  colnames(outlier)[i] <-colnames(DATA_nssamples)[i] 
  tmp.outlier <- vector()
  for (j in 1:10){
    tmp.iclust <- tmp.df[which(tmp.df$iClust==j), ]
    tmp.lm <- lm(tmp.iclust$NanoString ~ tmp.iclust$Illumina)
    tmp.iclust$CooksD <- cooks.distance(tmp.lm)
    tmp.outlier <- c(tmp.outlier, rownames(tmp.iclust)[which(tmp.iclust$CooksD>3*mean(tmp.iclust$CooksD))])
  }
  length(tmp.outlier) <- 100
  outlier[,i] <- tmp.outlier
  }

outlier <- data.frame(outlier[rowSums(is.na(outlier)) != ncol(outlier),])

write.csv(outlier, file="outlier.csv")

#we can also make another list which is to find out the ones that are incorrectly predicted, which genes are they outliers of
library(readr)
specimen_with_correlation <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Other Info/specimen_with_correlation.csv")

specimen_with_correlation$ID <- sub("-AD", "", specimen_with_correlation$MBID, fixed=TRUE)
specimen_with_correlation$genes <- NA

outlier_by_id <- list()

for (i in 1:nrow(specimen_with_correlation)){
    outlier_by_id[[i]] <- names(outlier)[which(outlier == specimen_with_correlation$ID[i], arr.ind=T)[, "col"]]
    names(outlier_by_id)[i] <- specimen_with_correlation$MBID[i]
    length(outlier_by_id[[i]]) <- 100
}

outliers_id <- data.frame(do.call(rbind, outlier_by_id))

write.csv(outliers_id, file="outlier_id.csv")
```

