---
title: "Selecting Candidate Genes from Metabric Samples using all 1980 samples"

output: pdf_document
---

Using all 1980 samples from original METABRIC dataset, select genes that are most diffrentially expressed in various iClusters


```{r}
library(readr)
library(limma)
library(randomForest)
```


```{r}
#loading data 

#turning Exp into a matrix
load("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Other Info/METABRIC_CURTIS_EXPRESSION.RData")
rownames(Exp) <- Exp$SYMBOL
#and trim down to only matrix
Exp <- Exp[, 3:ncol(Exp)]
Exp <- as.matrix(Exp)
Exp <- na.omit(Exp)
#now the metadata
IntClust <- read_csv("~/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Other Info/IntClust.csv")
IntClust$ID <- sub("-", ".", IntClust$ID, fixed=T)
#keeping only the ones that are in Exp
IntClust <- IntClust[which(IntClust$ID %in% colnames(Exp)), ]
#now order it the same way
IntClust <- IntClust[order(match(IntClust$ID, colnames(Exp))), ]


#now we have the right matrix and metadata files.  phew
save(Exp, file="mb_exp_matrix.RData")

load("mb_exp_matrix.Rdata")

```

Now doing the differential expression bit


```{r}
iclust_list <- list()
for (i in 1:10) {
  tmp <- IntClust
  tmp$IntClust <- as.character(tmp$IntClust)
  tmp$IntClust[which(tmp$IntClust != i)] = 0
  tmp$IntClust <- as.numeric(tmp$IntClust)
  iclust_list[[i]] <- tmp
  }  

```

We have the top 150 differentially expressed genes for each cluster
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
toptables <- list()
for (i in 1:10){
  design <- model.matrix(~ iclust_list[[i]]$IntClust)
  fit <- lmFit(Exp, design)
  fit <- eBayes(fit)
  test <- topTable(fit, number=150)
  toptables[[i]] <- test
}

#for (i in 1:10){
 #print(paste0("iClust", i, ":"))
 #print(rownames(toptables[[i]]))
  
#}

save(toptables, file="significantgenes.Rdata")
```


so let's make sure that they are all significant

```{r}
tmp_list <- list()
for (i in 1:10) {
  tmp_list[[i]] <- toptables[[i]][which(toptables[[i]]$P.Value<0.05), ]
  print(dim(tmp_list[[i]]))
    }

```


so now we have the list of genes, we can pick a list of genes and then we can now do the random forest....... 
let's keep 150 for IC1 and 2, and 50 genes for the rest

```{r}
genes <- c(rownames(toptables[[1]]), rownames(toptables[[2]]))

for (i in 3:10){
  genes <- c(genes, rownames(toptables[[i]])[1:50])
  }


#keeping unique ones only
genes <- unique(genes)

save(genes, file="genes for 1980 samples.RData")

```

now we try to trim down the data matrix, keeping only the genes that are useful

```{r}
mb_data_trimmed <- Exp[which(rownames(Exp) %in% genes), ]

dim(mb_data_trimmed)
```


now we have a matrix mb_data_trimmed, with rownames as important genes, and colnames are sample names

```{r}
#checking the colnames of the data matrix is in the same order as rows in the metadata df
sum(colnames(mb_data_trimmed)==IntClust$ID)
```

```{r}
#now testing if random forest can work with just one data.....
mb_data_trimmed <- t(mb_data_trimmed)
IntClust$IntClust <- as.factor(IntClust$IntClust)


rftest <- randomForest(mb_data_trimmed[-1,], IntClust$IntClust[-1])
predict(rftest, mb_data_trimmed[1,])

#now the random forest bit

rf_list_loocv <- lapply(1:nrow(mb_data_trimmed), function(i) {
  set.seed(100)
  tmp <- mb_data_trimmed[-i,]
  ic <- IntClust$IntClust[-i]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[i,], type = "response", predict.all=TRUE)
  return(data.frame(rownames(mb_data_trimmed)[i], tmp.pred$aggregate))
})


save(rf_list_loocv, file="rf_list_loocv.RData")

rf_loocv <- do.call(rbind.data.frame, rf_list_loocv)
sum(rf_loocv$rownames.mb_data_trimmed..i.==IntClust$ID)
#good now the rows are the same

rf_loocv_table <- table(rf_loocv$tmp.pred.aggregate, IntClust$IntClust)
save(rf_loocv_table, file="rf_loocv_table.RData")
diag(rf_loocv_table)/margin.table(rf_loocv_table, 2)

```

then we run 1 ic vs the rest for all the int clusters

```{r}
#start with iclust 1
IntClust$iClust1 <- as.character(IntClust$IntClust)
IntClust$iClust1[which(IntClust$iClust1 != 1)] = "others"
IntClust$iClust1 <- factor(IntClust$iClust1)




rf_list_ic1 <- lapply(1:nrow(mb_data_trimmed), function(i) {
  set.seed(100)
  tmp <- mb_data_trimmed[-i,]
  ic <- IntClust$iClust1[-i]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[i,], type = "response", predict.all=TRUE)
  return(data.frame(rownames(mb_data_trimmed)[i], tmp.pred$aggregate))
})

save(rf_list_ic1, file="rf_list_ic1.RData")


```


try with iCluster 2, using the kth fold thing (leave out 1:100, 101:200 etc)

```{r}
#first create a vector i - or a list???
list_i <- list()
for (j in 1:19) {
  list_i[[j]] <- seq(((j-1)*100+1),(j*100))
  #print(j)
}
list_i[[20]] <- seq(1901, 1980)

#then run it with a for loop to print i so we can find out if it's crashing ... 
#that's the plan anyway

#so create a list for result first
rf_ic2_kfcv_list <- list()
for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],]
  ic <- IntClust$iClust2[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],], type = "response", predict.all=TRUE)
  rf_ic2_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }

#and lets make it into a dataframe
rf_ic2_kfcv <- do.call(rbind.data.frame, rf_ic2_kfcv_list)

rf_ic2_kfcv_table<- table(rf_ic2_kfcv$tmp.pred.aggregate, IntClust$iClust2)
save(rf_ic2_kfcv_table, file="rf_ic2_ifcv_table.Rdata")
```

so let's try kth fold cross validation now, leaving 100 out at a time...

```{r}
#so create a list for result first
rf_kfcv_list <- list()
for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],]
  ic <- IntClust$IntClust[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],], type = "response", predict.all=TRUE)
  rf_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }

#and lets make it into a dataframe
rf_kfcv <- do.call(rbind.data.frame, rf_kfcv_list)

rf_kfcv_table<- table(rf_kfcv$tmp.pred.aggregate, IntClust$IntClust)
save(rf_kfcv_table, file="rf_kfcv_table.Rdata")
```

So now we use only the genes that are good for each group, and do individual group vs others.
For example, using the 150 genes that were selected for ic1, to classify ic1 vs others

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#so the files are: mb_data_trimmed, toptables for the genes, intclust for the ic information
#need list_i for k fold cross validation

#start with iC1
rf_selectedgenes_ic1_kfcv_list <- list()
for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[1]]))]
  ic <- IntClust$iClust1[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[1]]))], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic1_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }


rf_selectedgenes_ic1_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic1_kfcv_list)
rf_selectedgenes_ic1_kfcv_table<- table(rf_selectedgenes_ic1_kfcv$tmp.pred.aggregate, IntClust$iClust1)


#now iC2

rf_selectedgenes_ic2_kfcv_list <- list()
for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[2]]))]
  ic <- IntClust$iClust2[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[2]]))], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic2_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic2_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic2_kfcv_list)
rf_selectedgenes_ic2_kfcv_table<- table(rf_selectedgenes_ic2_kfcv$tmp.pred.aggregate, IntClust$iClust2)


#and ic3

rf_selectedgenes_ic3_kfcv_list <- list()
for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[3]])[1:50])]
  ic <- IntClust$iClust3[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[3]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic3_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic3_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic3_kfcv_list)
rf_selectedgenes_ic3_kfcv_table<- table(rf_selectedgenes_ic3_kfcv$tmp.pred.aggregate, IntClust$iClust3)
save(rf_selectedgenes_ic3_kfcv_table, file="rf_selectedgenes_ic3_kfcv_table.RData")

#ic4
rf_selectedgenes_ic4_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[4]])[1:50])]
  ic <- IntClust$iClust4[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[4]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic4_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic4_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic4_kfcv_list)
rf_selectedgenes_ic4_kfcv_table<- table(rf_selectedgenes_ic4_kfcv$tmp.pred.aggregate, IntClust$iClust4)
save(rf_selectedgenes_ic4_kfcv_table, file="rf_selectedgenes_ic4_kfcv_table.RData")

```


and we keep going to ic5 and beyond

```{r}
#ic5
rf_selectedgenes_ic5_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[5]])[1:50])]
  ic <- IntClust$iClust5[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[5]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic5_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic5_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic5_kfcv_list)
rf_selectedgenes_ic5_kfcv_table<- table(rf_selectedgenes_ic5_kfcv$tmp.pred.aggregate, IntClust$iClust5)
save(rf_selectedgenes_ic5_kfcv_table, file="rf_selectedgenes_ic5_kfcv_table.RData")



#ic6
rf_selectedgenes_ic6_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[6]])[1:50])]
  ic <- IntClust$iClust6[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[6]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic6_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic6_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic6_kfcv_list)
rf_selectedgenes_ic6_kfcv_table<- table(rf_selectedgenes_ic6_kfcv$tmp.pred.aggregate, IntClust$iClust6)
save(rf_selectedgenes_ic6_kfcv_table, file="rf_selectedgenes_ic6_kfcv_table.RData")



#ic7
rf_selectedgenes_ic7_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[7]])[1:50])]
  ic <- IntClust$iClust7[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[7]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic7_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic7_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic7_kfcv_list)
rf_selectedgenes_ic7_kfcv_table<- table(rf_selectedgenes_ic7_kfcv$tmp.pred.aggregate, IntClust$iClust7)
save(rf_selectedgenes_ic7_kfcv_table, file="rf_selectedgenes_ic7_kfcv_table.RData")

#ic8 (nearly there....)
rf_selectedgenes_ic8_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[8]])[1:50])]
  ic <- IntClust$iClust8[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[8]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic8_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic8_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic8_kfcv_list)
rf_selectedgenes_ic8_kfcv_table<- table(rf_selectedgenes_ic8_kfcv$tmp.pred.aggregate, IntClust$iClust8)
save(rf_selectedgenes_ic8_kfcv_table, file="rf_selectedgenes_ic8_kfcv_table.RData")


#ic9
rf_selectedgenes_ic9_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[9]])[1:50])]
  ic <- IntClust$iClust9[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[9]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic9_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic9_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic9_kfcv_list)
rf_selectedgenes_ic9_kfcv_table<- table(rf_selectedgenes_ic9_kfcv$tmp.pred.aggregate, IntClust$iClust9)
save(rf_selectedgenes_ic9_kfcv_table, file="rf_selectedgenes_ic9_kfcv_table.RData")
```
and finally


```{r}
#ic10
rf_selectedgenes_ic10_kfcv_list <- list()

for (k in 1:20){
  set.seed(100)
  tmp <- mb_data_trimmed[-list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[10]])[1:50])]
  ic <- IntClust$iClust10[-list_i[[k]]]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_data_trimmed[list_i[[k]],which(colnames(mb_data_trimmed) %in% rownames(toptables[[10]])[1:50])], type = "response", predict.all=TRUE)
  rf_selectedgenes_ic10_kfcv_list[[k]] <- data.frame(rownames(mb_data_trimmed)[list_i[[k]]], tmp.pred$aggregate)
  print(k)
  }
rf_selectedgenes_ic10_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic10_kfcv_list)
rf_selectedgenes_ic10_kfcv_table<- table(rf_selectedgenes_ic10_kfcv$tmp.pred.aggregate, IntClust$iClust10)
save(rf_selectedgenes_ic10_kfcv_table, file="rf_selectedgenes_ic10_kfcv_table.RData")
```

