---
title: "mb_997samples"
output: pdf_document
---

We will eventually go back to the original 997 discovery set to select genes.  

First we look at how


using 1980 samples
681 genes that has been selected
using ic10 package

```{r}
#load data
#mb_data_trimmed


features <- matchFeatures(Exp=t(mb_data_trimmed), Exp.by.feat="gene")
features <- normalizeFeatures(features, "scale")
res <- iC10(features)

res$class

sum(names(res$class)==IntClust$ID)

table(res$class, IntClust$IntClust)
prop.table(table(res$class, IntClust$IntClust), 2)
```

so now only get the "purest" discovey set samples to select genes

```{r}
#get the IDs
mb_discoveryset <- read.table("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Other Info/discovery_set_iC.txt", sep="\t", header = T)
mb_discoveryset$METABRIC_ID <-  sub("-", ".", mb_discoveryset$METABRIC_ID, fixed=T)

#load(Exp)

#now subset the datasets  to 977 samples
mb_977 <- Exp[, which(colnames(Exp) %in% mb_discoveryset$METABRIC_ID)]
IntClust977 <- IntClust[which(IntClust$ID %in% mb_discoveryset$METABRIC_ID), ]
```


creating list for each iclust in order to work out differential expression
```{r}
iclust_list <- list()
for (i in 1:10) {
  tmp <- IntClust977
  tmp$IntClust <- as.character(tmp$IntClust)
  tmp$IntClust[which(tmp$IntClust != i)] = 0
  tmp$IntClust <- as.numeric(tmp$IntClust)
  iclust_list[[i]] <- tmp
  } 
```


We have the top 150 differentially expressed genes for each cluster
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
toptables <- list()
for (i in 1:10){
  design <- model.matrix(~ iclust_list[[i]]$IntClust)
  fit <- lmFit(mb_977, design)
  fit <- eBayes(fit)
  test <- topTable(fit, number=150)
  toptables[[i]] <- test
}

# no number limit, but only significant ones
toptables_all <- list()
for (i in 1:10){
  design <- model.matrix(~ iclust_list[[i]]$IntClust)
  fit <- lmFit(mb_977, design)
  fit <- eBayes(fit)
  test <- topTable(fit, number=676)
  toptables_all[[i]] <- test[which(test$P.Value<0.05), ]
}

all_sig_genes <- vector()

for (i in 1:10){
  all_sig_genes <- c(all_sig_genes, rownames(toptables_all[[i]]))
}


```

so let's make sure that they are all significant

```{r}
tmp_list <- list()
for (i in 1:10) {
  tmp_list[[i]] <- toptables[[i]][which(toptables[[i]]$P.Value<0.05), ]
  print(dim(tmp_list[[i]]))
    }
```
so now we have the list of genes, we can pick a list of genes and then we can now do the random forest....... 
let's keep 150 for IC1 and 2, and 50 genes for the rest

```{r}
genes_977 <- c(rownames(toptables[[1]]), rownames(toptables[[2]]))

for (i in 3:10){
  genes_977 <- c(genes_977, rownames(toptables[[i]])[1:50])
  }


#keeping unique ones only
genes_977 <- unique(genes_977)

save(genes_977, file="genes for 977 samples.RData")

```


so let's now make a trimmed 977 samples with 676 genes dataframe

```{r}
mb_977_trimmedgenes <- mb_977[which(rownames(mb_977) %in% genes_977), ]
#so this is a 676 genes (rows) with 997 columns of samples

#and let's transform it and make sure the intclust 977 is also factors

mb_977_trimmedgenes <- t(mb_977_trimmedgenes)
#str(IntClust977)

#and check that the rownames and colnames ar ethe same
sum(rownames(mb_977_trimmedgenes)==IntClust977$ID)

```

and now the random forest with LOOCV 

```{r}
rf_977_loocv <- lapply(1:nrow(mb_977_trimmedgenes), function(i) {
  set.seed(100)
  tmp <- mb_977_trimmedgenes[-i,]
  ic <- IntClust977$IntClust[-i]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_977_trimmedgenes[i,], type = "response", predict.all=TRUE)
  return(data.frame(rownames(mb_977_trimmedgenes)[i], tmp.pred$aggregate))
})


save(rf_977_loocv, file="rf_977_loocv.RData")

rf_977_loocv_df <- do.call(rbind.data.frame, rf_977_loocv)

sum(rf_977_loocv_df$rownames.mb_977_trimmedgenes..i.==IntClust977$ID)

rf_977_loocv_table <- table(rf_977_loocv_df$tmp.pred.aggregate, IntClust977$IntClust)
save(rf_977_loocv_table, file="rf_977_loocv_table.RData")
```


individual cluster vs others and stick with LOOCV

```{r}
rf_977_676_ic1_kfcv_list <- list()
for (k in 1:977){
  set.seed(100)
  tmp <- mb_977_trimmedgenes[-k,]
  ic <- IntClust977$iClust1[-k]
  tmp.rf <- randomForest(tmp, ic)
  tmp.pred <- predict(tmp.rf, mb_977_trimmedgenes[k, ], type = "response", predict.all=TRUE)
  rf_977_676_ic1_kfcv_list[[k]] <- data.frame(rownames(mb_977_trimmedgenes)[[k]], tmp.pred$aggregate)
  print(k)
  }


rf_selectedgenes_ic1_kfcv <- do.call(rbind.data.frame, rf_selectedgenes_ic1_kfcv_list)
rf_selectedgenes_ic1_kfcv_table<- table(rf_selectedgenes_ic1_kfcv$tmp.pred.aggregate, IntClust$iClust1)



```

