---
title: "Random Forest"
output: pdf_document
---

Now by combining the trees together we have a forest



```{r include=FALSE}
library(lattice)
library(ggplot2)
library(caret)
library(doMC)
library(foreach)
library(iterators)
library(parallel)
load("nanoStringIC_DF.Rdata")
str(nanoStringIC_DF)
nanoStringIC_DF$iClust<-paste0("x",nanoStringIC_DF$iClust)
nanoStringIC_DF <- na.omit(nanoStringIC_DF)
rownames(nanoStringIC_DF)<-gsub("-","",rownames(nanoStringIC_DF))
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

```{r echo=FALSE}

ctrl <- trainControl(method = "repeatedcv", savePred=T, classProb=T, repeats= 5)
mod <- train(iClust~., data=nanoStringIC_DF, method = "rf", trControl = ctrl)

mod
```

