---
title: "Tweaking normalisation"
output: pdf_document
---


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(NanoStringNorm)
library(ggplot2)
library(reshape2)
#dev.off()

#save(mb_expression_matrix, file="mb_expression_matrix.RData")
load("mb_expression_matrix.RData")
load("nanoStringIC_DF.RData")

```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
batchNumber <- "Combined"
metadataFolder <- "/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Other Info/"
nanostringDataFolder <- paste0("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Digital Analyser/",batchNumber,"/")
outputFolder <- paste0("/Users/cope01/Documents/OneDrive - University Of Cambridge/Documents/PhD/Nanostring/Nanostring QC/",batchNumber,"QC/")

dir.create(outputFolder,showWarnings = FALSE)

iC10genesFile <- paste0(metadataFolder,"for_nanostring.txt")
metabricExpressionData <- paste0(metadataFolder,"METABRIC_CURTIS_EXPRESSION.RData")

#pdf(paste0(outputFolder,"metrics.pdf"), width=11, height=8)

nanoStringFiles <- list.files(nanostringDataFolder, pattern = "MB-AD", recursive = T, full.names = T)
nanoStringData <- lapply(nanoStringFiles, read.csv,
                         skip = 26, header = T, nrows = 773,quote=NULL,
                         stringsAsFactors = F)
#extract metabric id from file id
metabricID <- regmatches(nanoStringFiles, regexpr("MB-AD-[0-9]*", nanoStringFiles))

#rename count column to metabric ID
for (i in seq_along(nanoStringData)){
  colnames(nanoStringData[[i]])[colnames(nanoStringData[[i]])=="Count"] <-metabricID[[i]]
}

#convert list to dataframe
nanoStringData <- do.call("cbind", nanoStringData)
nanoStringData <- nanoStringData[, !duplicated(colnames(nanoStringData), fromLast = FALSE)] 

```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#Getting ready to correlate raw data with illumina data
nanoStringData_trimmed <- nanoStringData[, 4:ncol(nanoStringData)]
rownames(nanoStringData_trimmed) <- nanoStringData$Name
colnames(nanoStringData_trimmed) <- sub("-AD-", "-", colnames(nanoStringData_trimmed), fixed=TRUE)
nanoStringData_trimmed <- nanoStringData_trimmed[, match(colnames(mb_expression_matrix), colnames(nanoStringData_trimmed))]

mean(colnames(nanoStringData_trimmed)==colnames(mb_expression_matrix))

nanoStringData_trimmed <- nanoStringData_trimmed[which(rownames(nanoStringData_trimmed) %in% rownames(mb_expression_matrix)), ]

nanoStringData_trimmed <- nanoStringData_trimmed[match(rownames(mb_expression_matrix), rownames(nanoStringData_trimmed)), ]
mean(rownames(nanoStringData_trimmed)==rownames(mb_expression_matrix))

nanoStringData_trimmed <- as.matrix(nanoStringData_trimmed)
#now rows and columns are matched
nanoStringData_trimmed_ic <- merge(data.frame(t(nanoStringData_trimmed)), nanoStringIC_DF, by="row.names")

save(mb_expression_matrix, file="mb_expressioni_matrix.RData")
save(nanoStringData_trimmed, file="nanoStringData_trimmed.RData")
save(nanoStringData_trimmed_ic, file="nanoStringData_trimmed_ic.RData")


#see that the RSF1 gene is reasonably distributed
#which(rownames(mb_expression_matrix)=="RSF1")
plot(nanoStringData_trimmed[156, ],mb_expression_matrix[156, ])
plot(density(nanoStringData_trimmed["RSF1", ]))
ggplot(nanoStringData_trimmed_ic, aes(x=iClust, y=RSF1.x))+
  geom_boxplot()+
  theme_bw()

#it is... so we'll use this within each method
```

We have two sets of data for same samples and genes.  We now see which methods of normalisation should we use so that the nanostring data will correlate best with the illumina data.


Firstly we look at the correlation using raw count.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
#calculating the correlations
cors_gene_raw <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(nanoStringData_trimmed[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_raw <- cors_gene_raw[156, ]

#plotting the correlation
ggplot(cors_gene_raw, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(aes(x = rsf1_raw$cors, y = 0, xend = rsf1_raw$cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_raw, aes(x=rsf1_raw$cors, y=1.7, label=paste("RSF1,", round(rsf1_raw$cors, 2))))

```
```{r message=FALSE, warning=FALSE, include=FALSE}
#make a dataset that is normalisable
ns_raw <- nanoStringData

rownames(ns_raw) <- nanoStringData$Name
colnames(ns_raw) <- sub("-AD-", "-", colnames(ns_raw), fixed=TRUE)

ns_raw <- cbind(ns_raw[, 1:3], ns_raw[, match(colnames(mb_expression_matrix), colnames(ns_raw))])
mean(colnames(ns_raw)[4:ncol(ns_raw)]==colnames(mb_expression_matrix))

ns_raw <- rbind(ns_raw[750:773,], ns_raw[match(rownames(mb_expression_matrix), rownames(ns_raw)), ])
mean(rownames(ns_raw)[25:nrow(ns_raw)]==rownames(mb_expression_matrix))
```

Then try to normalise the data in the "standard" fashion
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised1<- NanoStringNorm(
  x = ns_raw,
  anno = NA,
  CodeCount = 'sum',
  Background = 'mean',
  SampleContent = 'housekeeping.sum',
  round.values = FALSE,
  take.log = FALSE,
  return.matrix.of.endogenous.probes = TRUE
)

#calculating the correlations
cors_genes_norm1 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(nsdata_normalised1[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_1 <- cors_genes_norm1[156, ]

#plotting the correlation
ggplot(cors_genes_norm1, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_1, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_1, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```

Now with z-value transformation method of normalisation:
```{r echo=FALSE, message=FALSE, warning=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised2<- NanoStringNorm(
  x = ns_raw,
  OtherNorm = "zscore",
  return.matrix.of.endogenous.probes = TRUE
)

#calculating the correlations
cors_genes_norm2 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(as.matrix(nsdata_normalised2)[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_2 <- cors_genes_norm2[156, ]

#plotting the correlation
ggplot(cors_genes_norm2, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_2, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_2, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```

Now we try the invrse normal transformation
```{r echo=FALSE, message=FALSE, warning=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised3<- NanoStringNorm(
  x = ns_raw,
  OtherNorm = "rank.normal",
  return.matrix.of.endogenous.probes = TRUE
)

#calculating the correlations
cors_genes_norm3 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(as.matrix(nsdata_normalised3)[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_3 <- cors_genes_norm3[156, ]

#plotting the correlation
ggplot(cors_genes_norm3, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_3, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_3, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```

We now use the quantile normalisation
```{r echo=FALSE, message=FALSE, warning=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised4<- NanoStringNorm(
  x = ns_raw,
  OtherNorm = "quantile",
  return.matrix.of.endogenous.probes = TRUE
)

#calculating the correlations
cors_genes_norm4 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(as.matrix(nsdata_normalised4)[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_4 <- cors_genes_norm4[156, ]

#plotting the correlation
ggplot(cors_genes_norm4, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_4, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_4, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```

We now try the variance stabilising normalisation.  There are 3 different ways of performing this.  
Firstly we fit and predict with "all" genes (controls and endogenous)
```{r echo=FALSE, message=FALSE, warning=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised5_1<- NanoStringNorm(
  x = ns_raw,
  OtherNorm = "vsn",
  return.matrix.of.endogenous.probes = FALSE,
  genes.to.fit = "all", 
  genes.to.predict = "all"
)

#calculating the correlations
cors_genes_norm5_1 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(as.matrix(nsdata_normalised5_1[[1]][15:231, 4:461])[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_5_1 <- cors_genes_norm5_1[156, ]

#plotting the correlation
ggplot(cors_genes_norm5_1, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_5_1, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_5_1, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```

We then try a method that fit the model on the "controls" and predict on the "endogenous".  
```{r echo=FALSE, message=FALSE, warning=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised5_2<- NanoStringNorm(
  x = ns_raw,
  OtherNorm = "vsn",
  return.matrix.of.endogenous.probes = FALSE,
  genes.to.fit = "controls", 
  genes.to.predict = "endogenous"
)

#calculating the correlations
cors_genes_norm5_2 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(as.matrix(nsdata_normalised5_2[[1]][15:231, 4:461])[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_5_2 <- cors_genes_norm5_2[156, ]

#plotting the correlation
ggplot(cors_genes_norm5_2, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_5_2, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_5_2, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```
Thirdly apply standard normalisation strategies as an alternative to the vsn affine transformation  So basically applies the glog2 variance stabilising transformation to the adjusted counts

```{r echo=FALSE, message=FALSE, warning=FALSE}
#normalisation using NanoStringNorm
nsdata_normalised5_3<- NanoStringNorm(
  x = ns_raw,
  CodeCount = "sum", 
  Background = "mean", 
  SampleContent = "top.geo.mean", 
  OtherNorm = "vsn",
  return.matrix.of.endogenous.probes = FALSE,
  genes.to.fit = "endogenous", 
  genes.to.predict = "endogenous", 
  calib="none"
)

#calculating the correlations
cors_genes_norm5_3 <- data.frame(
  cors=sapply(1:nrow(mb_expression_matrix),function(i){
    cor(as.matrix(nsdata_normalised5_3[[1]][15:231, 4:461])[i, ],mb_expression_matrix[i, ], method="spearman")
    }), 
  Genes = rownames(mb_expression_matrix)
  )

#create RSF1 subset
rsf1_5_3 <- cors_genes_norm5_3[156, ]

#plotting the correlation
ggplot(cors_genes_norm5_3, aes(x=cors))+
  geom_density()+
  theme_bw(base_size = 14)+
  xlab("R squared")+
  ggtitle("Correlation of Illumina and NanoString Raw Data")+
  geom_segment(data=rsf1_5_3, aes(x = cors, y = 0, xend = cors, yend = 1.5), linetype=2)+
  geom_text(data=rsf1_5_3, aes(x=cors, y=1.7, label=paste("RSF1,", round(cors, 3))))
```
```